Latent Variable Clustering for Classification of Events in the AT-TPC

### 2020-06-07 doc/Manuscript/article.tex
Updates from Overleaf
added table to mixae

{\color{cyan} [Ryan:] The commented-out paragraphs above I think could be better organized and condensed to eliminate some off-topic stuff. The outline below is an idea for how this section might tell its story.

\begin{itemize}
    \item Training high-performing neural networks from scratch often requires enormous datasets and computation time, particularly for vision tasks.
    \item However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks.
    \item For example, large neural networks which are trained on the ImageNet dataset --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes which are useful for numerous problems.
    \item Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large dataset).
    \item The training process then only has to fine-tune the network for the specific task. Since we are building on prior knowledge, learning becomes far more efficient.
    \item \citet{Kuchera2019} demonstrate the promise of this approach for classifying AT-TPC reaction products.
\end{itemize}

}

@@

In contrast with the classification work by \citet{Kuchera2019}, we do not append a classifying layer to the VGG16 network given that we do not have access to the ground truth labels at training time in the clustering regime. We instead consider the output from the convolutional blocks of the VGG16 network as a representation of the events on which we will apply a clustering algorithm.

@@

In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package. None of them provided any notable differences from the K-means results or were significantly worse. Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results. We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters. This is not the case for DBSCAN. 
\subsection{Autoencoder clustering}
\todo{look at grayscale images or colorbar with propto charge}

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/sim_mixae.pdf}
\caption[Clustering performance of MIXAE on simulated AT-TPC data]{Performance for the MIXAE model on simulated AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run. }\label{fig:mixae_sim}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/clean_mixae.pdf}
\caption[Clustering performance of MIXAE on filtered event AT-TPC data]{Performance for the MIXAE model on filtered AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_clean}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/real_mixae.pdf}
\caption[Clustering performance of MIXAE on full event AT-TPC data]{Performance for the MIXAE model on un-filtered AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_full}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{plots/clean_cluster_repr.pdf}
\caption[Selection of carbon events in differing clusters]{Selection of carbon events belonging to different clusters. Each row represents one of the clusters for the filtered data as shown in figure \ref{fig:mixae_confmat}. There seems to be no clear distinction between these rows.}\label{fig:filtered_mixae_clster_repr}	
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{plots/real_cluster_repr.pdf}	
\caption[Selection of proton events in differing clusters]{Selection of proton events belonging to different clusters. Each row represents one of the clusters for the full data as shown in figure \ref{fig:mixae_confmat}. We observe a tendency for the more noisy events to be placed in the "other" cluster}\label{fig:full_mixae_clster_repr}	
\end{figure}

>>>>

Training high-performing neural networks from scratch often requires enormous datasets and computation time. However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks. For example, large neural networks which are trained on the ImageNet dataset \cite{Russakovsky2015} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes which are useful for numerous problems. Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large dataset). The training process then only has to fine-tune the network for the specific task. Since we are building on prior knowledge in this case, learning becomes far more efficient and better performance can often be achieved. \citet{Kuchera2019} used machine learning methods to classify the products of $^{46}$Ar reactions in the AT-TPC, and they found that a CNN initialized with ImageNet weights resulted in the most successful classification.

@@

% In contrast with the classification work of \citet{Kuchera2019}, we do not append a classifying layer to the VGG16 network given that we do not have access to the ground truth labels at training time in the clustering regime. We instead consider the output from the convolutional blocks of the VGG16 network as a representation of the events on which we will apply a clustering algorithm.

% The clustering algorithm we found to give the strongest result is the K-means algorithm. This algorithm is contingent on having a distance measurement between two datapoints. We found that the standard choice of a euclidian norm on the vector differences provided the best outcomes.

In contrast with the classification work of \citet{Kuchera2019}, we do not have access to ground truth labels and are trying to solve a fundamentally different learning problem. Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events on which we will apply a clustering algorithm.

@@

In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package. None of them provided any notable differences from the K-means results or were significantly worse. Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results. We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters. This is not the case for DBSCAN.

\subsection{Autoencoder clustering}

\begin{table}[H]
\centering 
\caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data. In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.}\label{tab:clstr_vgg}
\input{./plots/mixae_clustering.tex}
\end{table}

@@

@@

@@

@@
